### 2015-09-12 paper/paper.tex

%\documentclass[preprint]{sigplanconf}

\documentclass[preprint]{llncs}

\usepackage{color}

\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
%% \usepackage{mathpartir}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}

\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}

% "define" code highlights for Java and Scala
\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,catch,class,def,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,val,var,while,with,yield},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=flexible,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numbers=none,                   % where to put the line-numbers
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
morekeywords={var},
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}

%%\usepackage{natbib}
%%\bibpunct();A{},
%%\let\cite=\citep

%include lhs2TeX.fmt
%include lhs2TeX.sty
%include forall.fmt

\pagestyle{plain}

%{\theorembodyfont{\sffamily} \newtheorem{theorem}{Theorem}}
%{\theorembodyfont{\sffamily} \newtheorem{lemma}{Lemma}}
%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{Lemma}
%\newenvironment{proof}{\textbf{Proof:\hspace{4mm}}}{$\Box$}
\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\yanlin[1]{\authornote{yanlin}{green}{#1}}
\newcommand{\hl}[1]{\textcolor{red}{#1}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\newcommand\mixin{\textbf{@Mixin}}

\begin{comment}
\author{Bruno C. d. S. Oliveira and Yanlin Wang}
\institute{The University of Hong Kong\\
\email{\{bruno,ylwang\}@cs.hku.hk}\\
\authorrunning{Bruno Oliveira and Yanlin Wang}} % abbreviated author list (for running head)
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Classless Java}
%%\subtitle{Solving The Expression Problem with Covariant Return Types}
%% Poor Man's Family Polymorphism

\maketitle

\begin{abstract}

\end{abstract}

%if False

%\category{D.3.2}{Programming Languages}
%                {Language Classifications}
%                [Functional Languages]
%\category{F.3.3}{Logics and Meanings of Programs}
%                {Studies of Program Constructs}
%                []

%\terms
%Languages
%
%\keywords
%Mixins, explicit effects, monads, aspect-oriented programming, parametricity,
%interference

%endif

%===============================================================================

\section{Introduction}\label{sec:intro}

\begin{itemize}

\item Using annotattions to implement a rich notion of traits with a 
mechanism to instantiate objects (the of method). Goal 1: is to reduce 
the amount of code that is required to program with interfaces and default 
methods. Goal 2: To provide a convenient means to do multiple inheritance 
in Java.

\item Specify the system more formally.

\item Show that we can model all trait operations

\item Implementation using Lombok

\item Case studies: The expression problem, Trivially
Case Studies from traits paper.

\end{itemize}

\section{Overview}\label{sec:ep}
Yanlin

* Explain what the Mixin annotations do using examples.

* Motivate the use of multiple inheritance in Java.

* Maybe use Marco's example (Point example).

We provide a Java annotation \mixin to provide default implementations for
various methods and a mechanism to instantiate objects. \mixin annotation helps
programmers to write less cumbersome code and instantiate interfaces in Java.

For example, interface \texttt{Point} annotated with \mixin:

\lstinputlisting[linerange=33-45]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT

\texttt{Point} has two (conceptually) member fields \texttt{X} and
\texttt{Y}. Methods \texttt{int X()} and \texttt{int Y()} serve as \emph{getter}
methods. Methods \texttt{void X(int X)} and \texttt{void Y(int Y)} serve as
\emph{setter} methods. Method \texttt{Point withX(int X)} updates field
\texttt{X} and returns \textbf{this}.

A typical and trivial implementation that programmers usually do is:

\lstinputlisting[linerange=82-112]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINTIMPL

\texttt{PointImpl} implements \texttt{Point} and provides a constructor with
quite mechanical code. What's worse, the implementation in \texttt{PointImpl}
may not be reused in a single inheritance language. However, with our approach,
the \mixin annotation will generate a static method \texttt{of} inside
\texttt{Point}. \texttt{of} makes use of Java anonymous classes and achieves the
same implementation as \texttt{PointImpl}.

\lstinputlisting[linerange=50-77]{../UseMixinLombok/src/test/TestLombok.java} % APPLY:linerange=POINT_OF

Inside the anonymous class in the annotated interface, the following code are
generated:
\begin{itemize}
\item For methods inside the interface with the form \texttt{Tx x()}:
  \begin{itemize}
   \item \texttt{x} is the getter method, with return type
     \texttt{Tx}. Conceptually, it is a member field with name \texttt{x} and
     type \texttt{Tx}.
   \item generate member field \texttt{\_x} of type \texttt{Tx}, initialized
     with \texttt{x}.
   \item generate implemented getter method:
       \begin{lstlisting}
           public Tx x() { return _x; }
       \end{lstlisting}
   \end{itemize}

\item For methods inside the interface with the form \texttt{void x(Tx x)}:
  \begin{itemize}
    \item check if exist method \texttt{Tx x()}. If not, generate error.
    \item generate implemented setter method:
       \begin{lstlisting}
           pubic void x(Tx x) { this.x = x; }
       \end{lstlisting}
    \end{itemize}

\item For methods with the form \texttt{T withX(Tx \_)}:
  \begin{itemize}
  \item if there is no \texttt{x} field, or type \texttt{Tx} does not match,
    then generate error.
  \item implement `withX` using the `of` method.
  \end{itemize}

\item For methods with the form of \texttt{T clone()}: Use \texttt{of} method as
  the constructor, to create a new object with the same field values as the
  current one.

\item For methods with the form of \texttt{T x(Tx \_)}:
  \begin{itemize}
    \item check if exist method \texttt{T x()}, if not, generate error.
    \item inside the inner class, generate
      \begin{lstlisting}
          pubic T x(Tx x) { this.x = x; return this;}
      \end{lstlisting}
  \end{itemize}
\end{itemize}

\section{Comparing to traits and mixins}
Haoyuan

   - vs both: we do automatic return type refinement, which has useful applications 
   (example: Expression Problem)

   - vs traits: we support of methods to create new objects (a replacement to constructors);
   Moreover we have the with and clone methods (we miss more applications for those). Show 
   how to model the operations on traits; discuss operations that we cannot model 
   (example: renaming).

   - vs mixins: we use the trait model of explicitly resolving conflicts. This is arguably 
   better for reasoning.

\section{Formal Semantics}\label{sec:typesafety}
Yanlin and Haoyuan

We need to show 2 things:

1) The dynamic semantics: what's the code that gets generated by a mixin annotation;

2) The type system: what programs to reject; properties: generation of type-safe/checkable code.

\bruno{The implementation is still missing the type system (rejecting some programs)!}

\section{Implementation}
Haoyuan

discuss implementation in lombok; and limitations.

\bruno{The implementation does not support separate compilation yet. Can we fix this?}

\section{Case studies}
Haoyuan and Yanlin

\subsection{A Trivial Solution to the Expression Problem}

\subsection{Other case studies} 
\bruno{The case studies still need to be implemented!}

Collections example from traits paper? 
(\yanlin{couldn't found source code})

Other case studies using multiple inheritance?

\section{Related Work}\label{sec:related}

- traits (original, variations, scala)
- mixins (original, scala)
- multiple inheritance 
- expression problem
- ...

%===============================================================================
\section{Conclusion}\label{sec:conclusion}

%===============================================================================

\bibliographystyle{splncs}
\bibliography{paper}

\appendix

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "."
%%% End: