### 2015-11-15 paper/sections/Implementation.tex
Minor changes to implementation section. I think it was already pretty good

\section{Implementation}
\begin{comment}
Haoyuan

discuss implementation in lombok; and limitations.

\bruno{The implementation does not support separate compilation yet. Can we fix this?}

\haoyuan{http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html}

\haoyuan{https://blog.frankel.ch/lombok-reduces-your-boilerplate-code}

\haoyuan{https://projectlombok.org/}
\end{comment}

One of the downside of Java is the code verbosity, e.g., getter and setter
methods for fields, unneeded type annotation, etc. Lombok project~\cite{} is a
Java tool that aims at removing (or reducing) these boilerplate code as far as
possible, via annotations. There are a number of annotations provided by the
original Lombok, including \textbf{@Getter}, \textbf{@Setter},
\textbf{@ToString} for generating getters, setters and \QM{toString} methods,
respectively.

Furthermore, Lombok provides a number of interfaces for users to create custom
transformations, as extensions to the original framework. By realizing these
transformations, with new annotations supported, users can generate new field
and method members, and inject them to current class types automatically at
compile time. More specifically, a transformation is based on a handler, which
acts on the AST from parsing the annotated node and returns a modified AST for
analysis and generation afterwards. Such a handler can either be a Javac handler
or an Eclipse handler.

Our implementation is based on extension to Lombok. In Eclipse, with an
interface annotated by \mixin, the automatic annotation processing is performed
and the information of the interface from compilation is captured in the
``Outline'' window, including all the methods inside the interface as well as
the generated ones.
 The custom transformation is easy and convenient to use.
 For example this means that the IDE functionality 
 content assist autocomplete will work for the newly generated methods.

The biggest reasons to use Lombok rather than using conventional Java annotation
processor is
\begin{itemize}
%this seamd like a commercial spot
%\item Lombok is byte-code based instead of source-code based, which makes client
%  code concise and easy to maintain. Such code generation is performed at
%  compile time to modify bytecode.
\item Lombock modify the generation process of the class files.
neither the source code is modified nor new  Java files are generated.
\item Moreover, and probably more importantly, Lombok is capable of generating
  code \emph{inside} a class/interface.This is the ability that conventional
  Java annotation processor do not provide.
\end{itemize}

Our prototype implementation using Lombok has certain limitation: 
\begin{itemize}
\item We do not support separate compilation yet at this
  stage, for in the implementation it is hard to capture a type declaration from
  its reference.
\item At this stage our implementation only realizes the Eclipse handler and our experiments are all conducted in Eclipse.
\end{itemize}