### 2015-11-18 paper/sections/Conclusion.tex
Minor changes in Introduction

\section{Conclusion}\label{sec:conclusion}

Before Java 8, concrete methods and static methods where not allowed
to appear in interfaces.  Java 8 allows static interface methods and
introduces \emph{default methods}, which allow for implementation
insides interfaces. This had an important positive consequence that
was probably overlooked by the Java design team: the concept of class
(in java) is now redundant and unneeded.  We define a subset of Java,
called ClassLess Java, where programs and (reusable) libraries can be
easily defined and used.  To avoid for some syntactic boilerplate
caused by Java not being originally designed to work without classes,
we introduce a new annotation:\mixin provide default implementations
for various methods (e.g. getters, setters, with-methods) and a
mechanism to instantiate objects. \mixin annotation helps programmers
to write less cumbersome code while coding in ClassLess Java; indeed
we think the obtained gain is so high that ClassLess Java with \mixin
annotation can be less cumbersome than full Java.\bruno{May need rewriting}