### 2015-11-25 paper/sections/RelatedWork.tex
Revised Related work (in progress)

\section{Related Work}\label{sec:related}
In this section we discuss related work and comparison to Classless Java.

%\yanlin{need to discuss: whether the comments for FTJ is fair.}

%No, was not fair, I know that work, it does modular type checking, and 
%more work after that they extend with more safe/modular typing things.
%also, Java8 IS a language extension on Java
%However, language extensions (including FTJ)
%have a natural drawback: the programmer has to learn new syntax. In contrast,
%our approach is completely compatible with the current Java language, so that
%programmers don't need to pay any learning cost to adapt to this new classless
%programming pattern. Another drawback which is particular for FTJ is that FTJ
%doesn't have type for traits, hence the correctness check of trait is done when
%type-checking classes. This choice makes the design of FTJ simpler but lost
%typechecking efficiency (one trait will be potentially checked multiple times if
%it is used in multiple classes).

\subsection{Multiple Inheritance in Object Oriented Languages}
Many authors have argued in favour or against multiple inheritance.  Multiple inheritance provides
expressive power, but it is difficult to model and implement, and can create
programs that are hard to reason about. These difficulties include the famous
diamond (fork-join) problem~\cite{bracha90mixin,Sak89dis}, conflicting methods, etc.
%and the yo-yo problem~\cite{taenzer1989problems}.  
To conciliate the need for expressive power and the need for simplicity, many
models have been proposed in over the years, including C++ virtual
inheritance, mixins~\cite{bracha90mixin} and traits~\cite{scharli03traits}.
They provide novel programming architecture models in the OO paradigm. In terms
of restrictions set on these models, C++ virtual inheritance aims at a more
general model, mixins added some restrictions on the model, and trait model is
the most restricted one (excluding states, instantiation, etc).
\bruno{How about Malayeri and Aldrich's paper? shouldn't that be discussed?}

\paragraph{C++ Approach.}
C++ tries to provide a general solution to multiple inheritance. Virtual
inheritance in C++ provides another solution to multiple inheritance (especially
the diamond problem by keeping only one copy of the base
class)~\cite{ellis1990annotated}, however suffers from object initialization
problem as pointed out by Malayeri et al.~\cite{malayeri2009cz}. It bypasses all
constructor calls to virtual superclasses, which would potentially cause serious
semantic errors. \bruno{What happens in our approach for the same case?}

\paragraph{Mixins.}
Mixins are a more restricted model than the C++ approach. Mixins allow to name
components that can be applied to various classes as reusable functionality
units. However, they suffer from linearisation: the order of mixin application
is relevant in often subtle and undesired ways. This hinders their usability
and the ability of resolving conflicts: the linearisation (total ordering) of
mixin inheritance cannot provide a satisfactory resolution in all cases and
restricts the flexibility of mixin composition. To fight those limitations, an
algebra of mixin operators is introduced~\cite{ancona2002calculus}, but this
raised the complexity of the approach, especially when constructors and fields
are considered~\cite{marco09FJigsaw}. Scala traits are in fact more like linearised mixins.
Scala avoids the object initialization
problem by disallowing constructor parameters, causing no ambiguity in cases
such as diamond problem. However this approach has limited expressiveness, and
suffers from all the problems of linearised mixin composition. Other languages, such as 
Python, also use linearized mixins.
\begin{comment}
Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
\Q@ class A use B,C {...new methods...}@\\*
can be expressed in python as \\*
\Q@ class Aux: ...new methods...@\\*
\Q@ class A(B,C,Aux): pass@ 
\end{comment}
\noindent Java interfaces and default methods does not use
linearisation: the semantics of Java \textbf{extends} clause in
interfaces is unordered and symmetric.

\bruno{rrevise latter?}
However, in pure Java, there is no mechanism for creating objects in
interfaces. Also, our approach supports proper constructor mechanism.

\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse (traits) and object factories
(classes). In this model, traits as units of reusable code, contain
only methods as reusable entities. Thus, no state and state
initialization are considered. Classes act as object factories,
requiring functionalities from multiple traits. Java 8 interfaces with
default methods are closely related to traits: concrete method
implementation are allowed (via the \textbf{default} keyword) inside
interfaces. The introduction of default methods opens the gate for
various flavours of multiple inheritance in Java, using interfaces.
Traits offer a trait algebra with operations like sum, alias and exclusion, provided
for explicit conflict resolution.  
Former work by Bono et al.~\cite{bono14}. provides details on
mimicking the trait algebra through Java 8 interfaces. 
We briefly recall the main points of their encoding;
however we propose a different representation of \textbf{exclusion}.
The author of ~\cite{bono14} agree that our revised version is
cleaner, typesafe and more direct.\bruno{do we want do say this?}

\newcommand\shortItem{\vspace{-1ex}}
\begin{itemize}
\item \textbf{Symmetric sum} can be obtained by simple multiple inheritance between interfaces.
\shortItem
    \begin{lstlisting}
    interface A { int x(); }    interface B { int y(); }    interface C extends A, B {}
    \end{lstlisting}
\shortItem
\item \textbf{Overriding} a conflict is obtained by specifying which super interface take precedence.
\shortItem
    \begin{lstlisting}
    interface A { default int m() {return 1;} } 
    interface B { default int m() {return 2;} }
    interface C extends A, B { default int m() {return B.super.m();} }
    \end{lstlisting}
\shortItem
\item \textbf{Alias} is creating  a new method delegating to the existing super interface.
\shortItem
    \begin{lstlisting}
    interface A { default int m() {return 1;} }
    interface B extends A { default int k() {return A.super.m();} }
    \end{lstlisting}
\shortItem

\item \textbf{Exclusion}: exclusion is also supported in Java, where method declarations can hide the default methods correspondingly in the super interfaces.
\shortItem
    \begin{lstlisting}
    interface A { default int m() {return 1;} }
    interface B extends A { int m(); }
    \end{lstlisting}
\shortItem
\end{itemize}

There are also proposals for extending Java (before Java8) with
traits. For example, FeatherTrait Java (FTJ)~\cite{Liquori08ftj} by
Liquori et al. extends the calculus of Featherweight Java
(FJ)~\cite{Igarashi01FJ} with statically-typed traits, adding
trait-based inheritance in Java.  Except for few, mostly syntactic
details, their work can be emulated/rephrased with Java8 interfaces.
There are also extensions to the original trait model, which have operations 
that default methods and interfaces cannot model, such as method renaming
(as in [Reppy2006]), which breaks structural subtyping.

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\bruno{who?} see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

%Nevertheless, we
%believe that in other languages the separation can be more practical
%than in Java, and we do not necessarally advocate that merging the two
%notions is a better approach in general.

%There are other limitations of our current approach, but they may be addressed
%in future work (see Section~\ref{sec:futurework}).

\begin{comment}
Besides, we support more features than the original trait model:
\begin{itemize}
\item We provide \texttt{of} methods for the annotated interfaces. During annotation processing time, the ``fields'' inside an interface are observed and a static method \texttt{of} is automatically injected to the interface with its arguments correspondingly. Such a method is a replacement to the constructors in original traits, making instantiation more convenient to use.
That is, in our approach there are only interfaces, our model requires a single concept, while the trait model requires traits \emph{and} classes.

\item We provide \texttt{with-} methods as auxiliary constructors. A \texttt{with-} method is generated for each field, just like record update, returning the new object with that field updated
%. A \texttt{clone} method is generated for the interface, returning a copy of the current object.
Furthermore, we do automatic return type refinement for these kind of methods. This feature is comparatively useful in big examples, making operations and behaviours more flexible.%, which we will demonstrate later.
\item We provide two options for generating setters. There are two kind of setters which are commonly used, namely \textit{void setters} and \textit{fluent setters}. The only difference is that a fluent setter returns the object itself after setting, thus supporting a pipeline of such operations. The generation depends on which type of setter is declared in the interface by users.
\end{itemize}
\end{comment}

% \begin{enumerate}
% \item Mixins allows to name components that can be applied to various classes as
%   reusable functionality units. However, they suffer from linearisation: the
%   order of mixin application is relevant in often subtitle and undesired
%   ways. constraints. This hinders their usability and their ability of resolving
%   conflicts: the linearisation (total ordering) of mixin inheritance cannot
%   provide a satisfactory resolution in all cases and restricts the flexibility
%   of mixin composition.  To fight those limitations, an algebra of mixin
%   operators is introduced~\cite{ancona2002calculus}, but this raised the
%   complexity of the approach, especially when constructors and fields are
%   considered~\cite{marco09FJigsaw}.  Our approach does not have the
%   linearisation problem, in that the semantics of Java \textbf{extends} clause
%   is unordered and symmetric.
% \item Simplifying the mixin algebras approach, traits draw a strong line between
%   units of reuse (traits) and object factories (classes) In this model,
%   traits~\cite{scharli03traits} units of reusable code, containing only methods
%   as reusable functionalities. Thus, no state/state initialization is
%   considered.

%   Classes act as object factories, requiring functionalities from multiple
%   traits.  Traits offers a trait algebra with operations like sum, alias and
%   exclusion, provided for explicit conflict resolution.

%   Concluding, (pure) traits do not allow state and they do not offers any reuse
%   instrument to ensure that state is coherently initialized when finally defined
%   in classes.  Traits can't be instantiated. This model requires two concepts
%   (traits and classes) to coexist and cooperate.

%   Some authors see this as good language design fostering good software
%   development by helping programmers to think about the structure of their
%   programs.  However, other authors see the need of two concepts and the absence
%   of state as drawbacks of this model. Our approach takes interfaces as units of
%   reuse, and meanwhile generates factory method for instantiation.

% \item C++ and Scala also try to provide solutions to multiple inheritance, but
%   both suffer from object initialization problems. Virtual inheritance in C++
%   provides another solution to multiple inheritance (especially the diamond
%   problem by keeping only one copy of the base class)~\cite{ellis1990annotated},
%   however suffers from object initialization problem as pointed out by Malayeri
%   et al.~\cite{malayeri2009cz}. It bypasses all constructor calls to virtual
%   superclasses, which would potentially cause serious semantic error. Scala
%   solution (very similar to linearised mixins, but misleadingly called traits in
%   the language) avoids this problem by disallowing constructor parameters,
%   causing no ambiguity in cases such as diamond problem.  This approach has
%   limited expressiveness, and suffers from all the problems of linearised mixin
%   composition.
%   Python also offers multiple inheritance via linearised mixins. Indeed in python any class is implicitly a mixin, and mixin composition informally expressed as\\*
%   \Q@ class A use B,C {...new methods...}@\\*
%   can be expressed in python as \\*
%   \Q@ class Aux: ...new methods...@\\*
%   \Q@ class A(B,C,Aux): pass@
%   Our approach not only does not involve linearised mixin problem, but also
%   support proper constructor mechanism.
% \end{enumerate}

% \input{sections/Compare.tex}

\subsection{ThisType/MyType/Extensibility}
%\marco{ this section is unfair. There are unsolved problems with ThisType in negative positions, namelly it breaks subtyping. The most common solution is to allow calling some methods only when the exact type is known. This demolish most advantages of interfaces.}

In certain situations, our approach allows automatic refinement for return
types. This is part of a bigger topic in class based languages: expressing and
preserving type recursion and (nominal/structural) subtyping at the same time.

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance. However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is huge limitation in class based object oriented
programming, and is exasperated by interface-only programming as we propose: no
type is ever exact since classes are not part of the language. A recent
article~\cite{Saito2013933} lights up this topic, since they propose two
features: exact statements \yanlin{do we support exact statements?}\marco{no we
  do not. When is that it may seams we do?} and non-inheritable methods; both
related to our work: any method generated inside of the \Q@of@ method is indeed
non-inheritable, since there is no class name to extends from, and exact
statements (a form of wild-card capture on the exact run-time type) could
capture the ``exact type'' of an object even in a class-less environment.

% The addition of MyType to a language will allow easy definition of
%binary methods, methods with recursive types (i.e., the same type of the
%receiver appears in the argument or return positions of methods), etc. MyType
%greatly enhances the expressiveness and extensibility of object-oriented
%programming languages.

%In our approach, we are using covariant-return types to simulate some uses of
%MyType. But our approach only works on method positive positions, whereas MyType
%is more general, as it works on any positions. Nevertheless our approach is
%still useful for modeling fluent interfaces and solving expression
%problems,etc.

\begin{comment}
\subsection{Type-Directed Translations/Syntactic Sugar}
\marco{I'm tring to merge this and the next one}
Language extensions are often implemented as syntactic sugar of the base
language. For example, Scala compiler supports XML syntax in normal Scala code
directly (after Scala ?, users need to import \texttt{scala-xml} library
manually). However, this approach is hard in terms of implementation, because it
requires extending the compiler. Also, this approach does not support combining
multiple extensions into one.

SugarJ~\cite{erdweg11sugarj} is a Java-based extensible programming language
that allows programmers to extend the base language with custom language
features by definitions in meta-DSLs (SDF, Stratego, etc). 
\yanlin{Is new syntax really a ``drawback''? I think for some system, like
  SugarJ, one of purpose IS to introduce these new syntax.}  Drawbacks: new
syntax. To create the extension, programmers have to work with multiple
languages (SDF, stratego, etc) while our approach works totally in Java
environment.

We can model certain types of language extensions with annotations 
only, but those extensions do not introduce new syntax: they 
merely do automatic code generation. 
\end{comment}
\subsection{Meta-programming Competes with Language extensions}
The most obvious solution to features addition to a language is language
extension.  It is often implemented as syntactic extensions that can be
desugared to the base language. For example, the Scala compiler was extended to
directly supports XML syntax. However, this approach does not support combining
multiple extensions into one. We are de facto creating a fork in the language,
and rarely the new fork gain enough traction to become the main language
release.  On this topic we mention SugarJ~\cite{erdweg11sugarj} - a Java-based
extensible language allowing programmers to extend it with custom features by
definitions in meta-DSLs (SDF, Stratego, etc).

On the other side, when the starting language have a flexible enough syntax and
a fast and powerful enough reflection mechanism, we may just need to play with
operator overloading and other language tricks to discover that the language
feature we need can be expressed as a simple library in our language. For
example, consider SQLAlchemy in python.

Java-like languages tend to sit in the middle of two extremes: libraries
can not influence the type system, so many solutions valid in python or other
languages could not be applicable, or may be applicable at the cost of loosing
safety.

Here (compile/load time) code generation comes at the rescue: if for a certain
feature (\mixin in our case) it is possible to use the original language syntax
to \emph{express-describe} any specific instantiation of such feature
(annotating a class and provide getters), then we can insert in the compilation
process a tool that exam and enrich the code before compilation. No need to
modify the original source; for example we can work on temporary files
\yanlin{Here what does temporary files mean? could explain more?}. Java is
a particular good candidate for this kind of manipulation since it already
provide ways to define and integrate such tools in its own compilation process:
in this way there is no need of temporary files, and there is a well defined way
of putting multiple extensions together.

Other languages offer even stronger support to safe code manipulation:
Template Haskell~\cite{sheard2002template}, F\# (type providers)~\footnote{http://research.microsoft.com/fsharp/} and MetaFjig (Active Libraries)~\cite{servetto2010metafjig}
all allow to execute code at compile time and to generate code/classes that are transparently integrated in the program that is being generated/processed/compiled.
In particular, MetaFjig offers a property called \emph{meta-level-soundness}. In short this property ensures by construction that library code (even if wrong or non nonsensical) would never generate ill-typed code. This is roughly equivalent to what we state and manually proof in Lemma 2 for our particular transformation.
Since MetaFjig is not working on annotated classes, there is no direct equivalence on the overall theorem of safety we shown.

\subsection{Formalization of Java8}
We provide a simple and well designed formalization for a subset of Java
including default/static interface methods and object initialization literals
(often called anonymous local inner classes).  A similar formalization was
drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize defender
(default) methods in Java. However this formalization is limited to model
exactly one method inside classes/interfaces.

% As an evidence of the attention and care present in our formalization work,
% while double-checking the behaviour of Java in side cases we have discovered a
% likely bug in the current \texttt{javac}.\marco{refer to before when we explain
%   the issue.}
\yanlin{bug claim removed.}