### 2015-11-26 paper/sections/CaseStudy.tex

\section{Case Studies}
In this section we conduct three case studies which reveals various
advantages using \mixin annotation. The first case study provides a
simple way to solve the Expression Problem, while supporting multiple,
independent extensions in Java.  The second case study show how to model an
embedded DSL for SQL languages with fluent interfaces. Finally, the third 
case study models a simple game, and compares our implementation with an 
existing one, showing that the amount of code is reduced significantly using \mixin.

\subsection{A Trivial Solution to the Expression Problem with Object Interfaces}

The \emph{Expression Problem} (EP)~\cite{wadler98expression}
is a well-known problem about modular extensibility issues in software
evolution. Recently, a new solution~\cite{eptrivially} that uses only
covariant type refinement that has been shown to work in Java. When
that solution is modelled with interfaces and default methods, it can
even provide independent extensibility: that is, the ability to
assemble a system, from multiple, independently developed extensions.
Unfortunatelly, a plain Java solution has alot of code verbosity in
Java due to code required for instantiation, which makes the solution 
cumbersome to use. The \mixin annotation helps to remove such code, and 
makes the solution quite easy to use.
% \bruno{We should try to do some space saving in here. You could, for example 
% combine the code for the initial system with the code for a new operation in a figure/table 
% with 2 colums. Then just refer to the Table/Figure in the text.}

\paragraph{Initial System}
In the formulation of the EP there is an initial system that models
arithmetic expressions with only literals and addition, and an initial 
operation:

\begin{lstlisting}[multicols=2]
interface Exp { int eval(); }
@Obj interface Lit extends Exp {
	int x();
	default int eval() { return x(); }
}
@Obj interface Add extends Exp {
	Exp e1(); Exp e2();
	default int eval() {
		return e1().eval() + e2().eval();
	}
}
\end{lstlisting}

% \begin{multicols}{2}
% \lstinputlisting[linerange=34-44]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
% \end{multicols}

\noindent \texttt{Exp} is the common super-interface with an evaluation operation
\texttt{eval()} inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend
interface \texttt{Exp} with default implementations for the \texttt{eval}
operation. The number field \texttt{x} of a literal is represented as a getter
method \texttt{x()} and expression fields (\texttt{e1} and \texttt{e2}) of an addition as
getter methods \texttt{e1()} and \texttt{e2()}.

\paragraph{Adding a New Type of Expressions}
It is easy to add new types of expressions to the code in the initial system in a
modular way. For example, the following code shows how to add subtraction.

\lstinputlisting[linerange=48-49]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB

\paragraph{Adding a New Operation} The difficulty of the EP in object-oriented 
languages arizes from adding new operations. For example, adding a pretty printing 
operation, would typically involve changing all of the existing code. However, a solution 
to the EP forbids this and it also forbids using casts. In other words, it should be possible 
to add the pretty printing operation in a type-safe and modular way.  
It turns out that, this can be done easily, using our object interfaces approach. 
The following code shows how to add the new operation \texttt{print}.

\lstinputlisting[linerange=53-61]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
%\bruno{Please add a Java comment to the code emphasizing that the types of the getters are refined!}
%\bruno{please indent code properly (the getters are not aligned with default))}

\noindent  The interface \texttt{ExpP} extending interface \texttt{Exp} is defined with the
extra method \texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are
defined with default implementations of \texttt{print()}, extending
base interfaces \texttt{Lit} and \texttt{Add}, respectively. Importantly, note
that in \texttt{AddP}, the types of the ``\emph{fields}'' (i.e. the getter methods) \texttt{e1} 
and \texttt{e2} are refined. If the types were not refined then the body of the \texttt{print}
method in \texttt{AddP} would fail to type-check.

\paragraph{Independent Extensibility}
To show that our approach supports independent extensibility~\cite{zenger05independentlyextensible}, we first define a
new operation \texttt{collectLit} (which collects all
literal components in an expression) on expressions. For space reasons, 
we omit the definitions of the methods:

\begin{lstlisting}[]
interface ExpC extends Exp { List<Integer> collectLit(); }
@Obj interface LitC extends Lit, ExpC {
    default List<Integer> collectLit() { ... }}
@Obj interface AddC extends Add, ExpC {
    ExpC e1(); ExpC e2(); //return type refined!
    default List<Integer> collectLit() { ... }}
\end{lstlisting}

\noindent Now we combine the two extensions (\texttt{print} and
\texttt{collectLit}) together:

\lstinputlisting[linerange=85-87]{../UseMixinLombok/src/test/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY

\noindent \texttt{ExpPC} is the new expression interface supporting
\texttt{print} and \texttt{collectLit} operations; \texttt{LitPC} and
\texttt{AddPC} are the extended variants. Notice that except for the routine of
\textbf{extends} clauses, no glue code is required. Return types of
\texttt{e1,e2} are also automatically refined to \texttt{ExpPC}.

Note that the code for instantiation is automatically generated by \mixin. 
So, for example creating a simple expression of type \texttt{ExpPC} is 
as simple as:
\begin{lstlisting}
ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
\end{lstlisting}
\noindent In contrast, in a pure Java solution, the tedious instantiation code would need 
to be defined manually.

\subsection{Embedded DSLs with Fluent Interfaces}\label{sec:dsls}
Since the style of fluent interfaces was invented in Smalltalk as method
cascading, more and more languages came to support fluent interfaces, including
JavaScript, Java, C++, D, Ruby, Scala, etc. For most languages, to create fluent
interfaces, programmers have to either handwrite everything or create a wrapper
around the original non-fluent interfaces using \textbf{this}. In Java, there
are several libraries (including jOOQ, op4j, fluflu, JaQue, etc) providing useful
fluent APIs. However most of them only provide a fixed set of predefined fluent
interfaces. Fluflu enables the creation of a fluent API and implements control
over method chaining by using Java annotations. However methods that returns
\textbf{this} are still hand-written.

The \mixin annotation can also be used to create fluent interfaces.  When
creating fluent interfaces with \mixin, there are two main advantages:
\begin{enumerate}
\item Instead of forcing programmers to hand write code using \textbf{return
    this}, our approach with \mixin annotation removes this verbosity and
  automatically generate fluent setters.
\item Along the extension direction, the return types of fluent setters are
  automatically refined.
\end{enumerate}

\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
The first query language \texttt{Database}  models
\texttt{select}, \texttt{from} and \texttt{where} clauses:
\lstinputlisting[linerange=6-11]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
% \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward 
% to have to use the of explicitly to initialized all fields.}

\noindent The main benefit that fluent methods give
us is the convinience of method chaining:

\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1

\noindent Note how all the logic for the fluent setters is automatically provided by the \mixin annotation.

\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
\texttt{orderBy} which orders the result records by a field that users
specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
feature, and the return type of fluent setters in
\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
\lstinputlisting[linerange=15-18]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
This way, when a \texttt{ExtendedDatabase} query created, 
all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling 
\texttt{orderBy}.

\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/jooq/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2

\subsection{A Maze Game}
The last case study is a simplified variant of Maze game, which is often
used~\cite{gof,bono14} 
to evaluate code reuse ability related to inheritance and design
patterns. In the game, there is a player with the goal of collecting
as many coins as possible. She may enter a room with several doors to
be chosen among. This is a good example because it involves code reuse
(different kinds of doors inherit a common type, with different
features and behaviour), multiple inheritance (a special kind of door
may require features from two other door types) and it also shows how
to model operations \texttt{symmetric sum}, \texttt{override} and
\texttt{alias} as trait-oriented programming. The game has been
implemented using plain Java 8 and default methods by Bono
et. al~\cite{bono14}, and the code for that implementation is
available online. We reimplemented the game using \mixin. Due to space
constraints, we omit the code here. The following table summarizes
the number of lines of code and classes/interfaces in each implementation:

\vspace{5pt}
\hspace{80pt}\begin{tabular}{ccc}
\hline
            & SLOC   & \# of classes/interfaces \\ \hline
Bono et al. & 335    & 14                       \\
Ours        & 199    & 11                       \\
\rowcolor[HTML]{C0C0C0} 
Reduced by  & 40.6\% & 21.4\%                   \\ \hline
\end{tabular}
\vspace{5pt}

\noindent The \mixin annotation allowed us to reduce the interfaces/classes used
in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
reductions was due to the replacement of instantiation classes with
generated \texttt{of} methods. The number of source lines of code (SLOC)
was reduced by 40\% due to both the removal of instantiation overhead and
generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}. 
To ensure a fair comparison, we used the same coding style as Bono et al.'s.