### 2016-01-26 OOPSLA16/paper.tex
OOPSLA template.

\documentclass[10pt,preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.
% nocopyrightspace ...

\usepackage{amsmath,amssymb}
\usepackage{balance,bbm,booktabs}
\usepackage{color}
\usepackage{etex}
\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{mathpartir,mathptmx,mdwlist,morefloats,multirow}
\usepackage{pstricks,pst-node,pst-tree,pgf}
\usepackage{rotating}
\usepackage{stmaryrd}
\usepackage{theorem}
\usepackage{url}
\usepackage{verbatim}
\usepackage{wrapfig}
\usepackage{xspace}

\usepackage[normalem]{ulem}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{dirtree}

\newcommand{\Or}{\ |\ }

\newcommand{\cL}{{\cal L}}
\input{predef/def.tex}
\input{predef/java.tex}
\hyphenation{}

\pagestyle{plain}

\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Title Text}
\subtitle{Subtitle Text, if any}

\authorinfo{Name1}
           {Affiliation1}
           {Email1}
\authorinfo{Name2\and Name3}
           {Affiliation2/3}
           {Email2/3}

\maketitle

\begin{abstract}
This is the text of the abstract.
\end{abstract}

\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2

\input{sections/Introduction.tex}

\input{sections/Formalization.tex}

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}

\end{document}