### 2016-01-26 OOPSLA16/sections/Formalization.tex

\section{Formal Semantics}\label{sec:formalization}

\begin{figure*}[t]
\[\begin{array}{llll}
T & ::= & I_1.\cdots.I_n                                                                                                         & \hspace{1in} \textrm{path} \\
e & ::= & x \Or e.m(\overline{e}) \Or T.m(\overline{e}) \Or T.\mathbf{super}.m(\overline{e}) \Or x = e;e' \Or obj         & \hspace{1in} \textrm{expressions} \\
obj & ::= & \mathbf{new}\ T()\ \{\overline{field}\ mh_1\{\mathbf{return}\ e_1;\}\cdots mh_n\{\mathbf{return}\ e_n;\}\}& \hspace{1in} \textrm{object creation} \\
field & ::= & T\ f=x;                                                                                              & \hspace{1in}  \textrm{field declaration} \\
\mathbb{I} & ::= & ann\ \mathbf{interface}\ I\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}& \hspace{1in} \textrm{interface declaration} \\
meth & ::= & \mathbf{static}\ mh\{\mathbf{return}\ e;\} \Or \mathbf{default}\ mh\{\mathbf{return}\ e;\} \Or mh; & \hspace{1in} \textrm{method declaration} \\
mh & ::= & T_0\ m(T_1\ x_1\cdots T_n\ x_n)  & \hspace{1in} \textrm{method header} \\
ann & ::= & \mathbf{@Family} \Or \emptyset & \hspace{1in} \textrm{annotations} \\
\Gamma & ::= & x_1:T_1\cdots x_n:T_n & \hspace{1in} \textrm{environment}
\end{array}\]
\caption{Grammar.}
\end{figure*}

\begin{figure*}[t]
\[\begin{array}{lll}
Exact(T_1.I,\ T_2) & = & Exact(I,\ Exact(T_1,\ T_2)) \\
Exact(I,\ I_1.\cdots.I_n) & = & Exact2(I,\ I_1.\cdots.I_i) \\
& & \ \ \ \ \textrm{for the smallest }i\in[0,n],\forall j\in(i,n],Exact2(I,\ I_1.\cdots.I_j) = \emptyset \\
Exact2(I,\ \emptyset) & = & CT(I) \\
Exact2(I,\ T_0.I_0) & = & \mathbb{I} \\
& & \ \ \ \ \textrm{if } \mathbb{I}\in\overline{\mathbb{I}} \textrm{ and } name(\mathbb{I})=I \\
& & Exact(I,\ Exact(T,\ T_0)) \\
& & \ \ \ \ \textrm{if } T\in\overline{T} \textrm{ and }\forall T'\in\overline{T}, Exact(I,\ Exact(T',\ T_0))\in\{\ Exact(I,\ Exact(T,\ T_0)),\ \emptyset\ \} \\
& & \textrm{where } CT(T_0.I_0) = ann\ \mathbf{interface}\ I_0\ \mathbf{extends}\ \overline{T}\ \{\overline{meth}\ \overline{\mathbb{I}}\}
\end{array}\]
\caption{Typing rules.}
\end{figure*}