### 2016-03-23 OOPSLA16/sections/CaseStudy.tex

\section{Case Studies}
In this section we conduct \numOfCaseStudies case studies which reveal various
advantages of using the \mixin annotation. The first case study provides a
simple way to solve the Expression Problem while supporting multiple,
independent extensions in Java.  The second case study shows how to model an
embedded DSL for SQL languages with fluent interfaces. Finally, the third
case study models a simple game, and compares our implementation with an
existing one, showing that the amount of code is reduced significantly using \mixin.

\subsection{A Trivial Solution to the Expression Problem with Object Interfaces}

The \emph{Expression Problem} (EP)~\cite{wadler98expression} is a well-known
problem about modular extensibility issues in software evolution. Recently, a
new solution~\cite{eptrivially} using only covariant type refinement was
proposed. When this solution is modeled with interfaces and default methods, it
can even provide independent extensibility~\cite{zenger05independentlyextensible}: the ability to assemble a system
from multiple, independently developed extensions. Unfortunately, the
required instantiation code makes a plain Java solution verbose and cumbersome
to use. The \mixin annotation is enough to remove the boilerplate code, making
the presented approach very appealing.

\paragraph{Initial System}
In the formulation of the EP, there is an initial system that models
arithmetic expressions with only literals and addition, and an initial
operation \texttt{eval} for expression evaluation.
As shown in Figure~\ref{fig:ep}, \texttt{Exp} is the common
super-interface with operation \texttt{eval()}
inside. Sub-interfaces \texttt{Lit} and \texttt{Add} extend interface
\texttt{Exp} with default implementations for the \texttt{eval} operation. The
number field \texttt{x} of a literal is represented as a getter method
\texttt{x()} and expression fields (\texttt{e1} and \texttt{e2}) of an addition
as getter methods \texttt{e1()} and \texttt{e2()}.

\begin{figure*}
\centering
\saveSpaceFig
\begin{tabular}{l|l}
\lstinputlisting[linerange=34-44]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_INIT
&
\lstinputlisting[linerange=55-65]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_PRINT
\end{tabular}
\caption{The Expression Problem (left: initial system, right: code for adding
  print operation).}\label{fig:ep}
\end{figure*}% \bruno{Yanlin: code is overflowing the margins. Note that
  % there are 2 spare lines that can be used on the right!}

\paragraph{Adding a New Type of Expressions}
In the OO paradigm, it is easy to add new types of expressions. For example, the
following code shows how to add subtraction.

\lstinputlisting[linerange=48-51]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=EXPRESSION_SUB

%\marco{text under need to be improved}
\paragraph{Adding a New Operation} The difficulty of the EP in OO
languages arises from adding new operations. For example, adding a pretty printing
operation would typically change all existing code. However, a solution
%to the EP forbids this and it also forbids using casts. In other words, it should be possible
should add operations in a type-safe and modular way. This
turns out to be easily achieved with the assistance of \mixin.  The code in
Figure~\ref{fig:ep} (on the right) shows how to add the new operation \texttt{print}.
Interface \texttt{ExpP} extends \texttt{Exp} with the extra method
\texttt{print()}. Interfaces \texttt{LitP} and \texttt{AddP} are defined with
default implementations of \texttt{print()}, extending base interfaces
\texttt{Lit} and \texttt{Add}, respectively. Importantly, note that in
\texttt{AddP}, the types of ``\emph{fields}'' (i.e. the getter methods)
\texttt{e1} and \texttt{e2} are refined. If the types were not refined then
the \texttt{print()} method in \texttt{AddP} would fail to type-check.

\paragraph{Independent Extensibility}
To show that our approach supports independent extensibility, we first define a
new operation \texttt{collectLit}, which collects all
literal components in an expression. For space reasons,
we omit the definitions of the methods:

\begin{lstlisting}[]
interface ExpC extends Exp {
    List<Integer> collectLit(); }
@Obj interface LitC extends Lit, ExpC {...}
@Obj interface AddC extends Add, ExpC {
    ExpC e1(); ExpC e2(); ...}
\end{lstlisting}

\noindent Now we combine the two extensions (\texttt{print} and
\texttt{collectLit}) together:

\lstinputlisting[linerange=89-91]{../UseMixinLombok/src/casestudy/ep/TestExpression.java}% APPLY:linerange=INDEPENDENT_EXTENSIBILITY

\noindent \texttt{ExpPC} is the new expression interface supporting
\texttt{print} and \texttt{collectLit} operations; \texttt{LitPC} and
\texttt{AddPC} are the extended variants. Notice that except for the routine of
\textbf{extends} clauses, no glue code is required. Return types of
\texttt{e1,e2} must be refined to \texttt{ExpPC}.

Note that the code for instantiation is automatically generated by \mixin.
Creating a simple expression of type \texttt{ExpPC} is
as simple as:
\begin{lstlisting}
ExpPC e8 = AddPC.of(LitPC.of(3), LitPC.of(4));
\end{lstlisting}
\noindent Without our approach, tedious instantiation code would need
to be defined manually.

\subsection{Embedded DSLs with Fluent Interfaces}\label{sec:dsls}
Since the style of fluent interfaces was invented in Smalltalk as method
cascading, more and more languages came to support fluent interfaces, including
JavaScript, Java, C++, D, Ruby, Scala, etc. In most languages, to create fluent
interfaces, programmers have to either hand-write everything or create a wrapper
around the original non-fluent interfaces using \textbf{this}. In Java, there
are several libraries (including jOOQ, op4j, fluflu, JaQue, etc) providing useful
fluent APIs. However most of them only provide a fixed set of predefined fluent
interfaces. Fluflu enables the creation of a fluent API and implements control
over method chaining by using Java annotations. However methods that returns
\textbf{this} are still hand-written.

The \mixin annotation can also be used to create fluent interfaces.  When
creating fluent interfaces with \mixin, there are two main advantages:
\begin{enumerate}
\item Instead of forcing programmers to hand write code using \textbf{return
    this}, our approach with \mixin annotation removes this verbosity and
  automatically generates fluent setters.
\item The approach supports extensibility: the return types of fluent setters are
  automatically refined.
\end{enumerate}

\noindent We use embedded DSLs of two simple SQL query languages to illustrate.
The first query language \texttt{Database}  models
\texttt{select}, \texttt{from} and \texttt{where} clauses:
\lstinputlisting[linerange=6-10]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE
% \bruno{Wouldn't it be better to have a static method create() that calls of("","","")) and initializes all the fields? It seems awkward
% to have to use the of explicitly to initialized all fields.}

\noindent The main benefit that fluent methods give
us is the convenience of method chaining:

\lstinputlisting[linerange=33-33]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY1

\noindent Note how all the logic for the fluent setters is automatically provided by the \mixin annotation.

\paragraph{Extending the Query Language} The previous query language can be extended with a new feature
\texttt{orderBy} which orders the result records by a field that users
specify. With \mixin programmers just need to extend the interface \texttt{Database} with new
features, and the return type of fluent setters in
\texttt{Database} is automatically refined to \texttt{ExtendedDatabase}:
\lstinputlisting[linerange=14-18]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_DATABASE_EXT
In this way, when a query created using \texttt{ExtendedDatabase},
all the fluent setters return the correct type, and not the old \texttt{Database} type, which would prevent calling
\texttt{orderBy}.

\lstinputlisting[linerange=39-39]{../UseMixinLombok/src/casestudy/sql/DatabaseTest.java}% APPLY:linerange=FLUENT_QUERY2

\subsection{A Maze Game}
This case study is a simplified variant of a Maze game, which is often
used~\cite{gof,bono14}
to evaluate code reuse ability related to inheritance and design
patterns. In the game, there is a player with the goal of collecting
as many coins as possible. She may enter a room with several doors to
be chosen among. This is a good example because it involves code reuse
(different kinds of doors inherit a common type, with different
features and behavior), multiple inheritance (a special kind of door
may require features from two other door types) and it also shows how
to model operations \texttt{symmetric sum}, \texttt{override} and
\texttt{alias} from trait-oriented programming. The game has been
implemented using plain Java 8 and default methods by Bono
et. al~\cite{bono14}, and the code for that implementation is
available online. We reimplemented the game using \mixin. Due to space
constraints, we omit the code here. The following table summarizes
the number of lines of code and classes/interfaces in each implementation:

\vspace{5pt}
\begin{tabular}{ccc}
\hline
            & SLOC   & \# of classes/interfaces \\ \hline
Bono et al. & 335    & 14                       \\
Ours        & 199    & 11                       \\
\rowcolor[HTML]{C0C0C0}
Reduced by  & 40.6\% & 21.4\%                   \\ \hline
\end{tabular}
\vspace{5pt}

\noindent The \mixin annotation allowed us to reduce the interfaces/classes used
in Bono et al.'s implementation by 21.4\% (from 14 to 11). The
reductions were due to the replacement of instantiation classes with
generated \texttt{of} methods. The number of source lines of code (SLOC)
was reduced by 40\% due to both the removal of instantiation overhead and
generation of getters/setters. %\bruno{explain why? we generate code for getters/setters?}.
To ensure a fair comparison, we used the same coding style as Bono et al.'s.

\subsection{Refactoring a Compiler}
The last case study applies our approach to a slightly larger code base of a simple
one-pass compiler for \lstinline{C0} (a subset of \lstinline{C}). This compiler
is a handwritten, monolithic compiler used for educational purposes at Aarhus
University, Denmark\footnote{Source code available at
  http://cs.au.dk/~mis/dOvs/Czero.java}. \lstinline{C0} is restricted to
integers, several control structures and function declaration/definition and
basic I/O statements. All the code (including parsing, code generation for
various structures, error handling) is entangled in one big file. Following our
approach with \mixin, we refactored the code to several smaller interfaces
(including \lstinline{Constants}, \lstinline{MemberFields}, interfaces for
parsing different structures, etc). After the refactoring, the new code does not
reduce code as previous case studies. SLOC for original code without comments
and blank lines is 828, while ours is 830 (for fair comparison, we put all
interfaces into one file when calculating SLOC). However, the case study shows
two things: Classless Java can be applied to real code base such as compilers;
code becomes more modular without the sacrifice of code amount/simplicity.
\marco{Why we do not cite the result of the former study? the one with object algebraes
where the number of lines grown quite a while?}