### 2016-03-23 OOPSLA16/paper.tex
abstract

\documentclass[10pt,preprint,nocopyrightspace,numbers]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.
% nocopyrightspace ...

\usepackage{listings}
\usepackage{multicol}
\usepackage{microtype}%if unwanted, comment out or use option "draft"
\usepackage[table,xcdraw]{xcolor}
\usepackage{color}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}

\usepackage{listings}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[normalem]{ulem}

\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{mathpartir}
\usepackage{xspace}

\input{predef/def.tex}
\input{predef/java.tex}

\newcommand{\Or}{\ |\ }

\newcommand{\cL}{{\cal L}}
\hyphenation{}

\pagestyle{plain}

\newcommand{\authornote}[3]{{\color{#2} {\sc #1}: #3}}
\newcommand{\authorText}[2]{{\color{#1}#2}}
\newcommand\bruno[1]{\authornote{bruno}{red}{#1}}
\newcommand\yanlin[1]{\authornote{yanlin}{purple}{#1}}
\newcommand\marco[1]{\authornote{marco}{blue}{#1}}
\newcommand\haoyuan[1]{\authornote{haoyuan}{cyan}{#1}}
\newcommand\brunoT[1]{\authorText{red}{#1}}
\newcommand\yanlinT[1]{\authorText{purple}{#1}}
\newcommand\marcoT[1]{\authorText{blue}{#1}}
\newcommand\haoyuanT[1]{\authorText{cyan}{#1}}

\newcommand\delete[1]{\textcolor{red}{\sout{#1}}}

\newcommand\sem[1]{\llbracket #1 \rrbracket_r}
\newcommand\sems[1]{\llbracket #1 \rrbracket_s}
\newcommand\tsem[1]{\llbracket #1 \rrbracket}
\newcommand{\rbm}[1]{\raisebox{-2.0ex}[0.5ex]{#1}}
\newcommand\nat[0]{\mathbb{N}}
\newcommand\unit[0]{\mathbbm{1}}

\newcommand\mixin{\mixinAnn\xspace}
%\renewcommand{\paragraph}[1]{\vspace{5pt}\noindent{\bf #1}}
\newenvironment{listing}{\vspace{-3pt}\begin{lstlisting}}{\end{lstlisting}\vspace{-3pt}}

%\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{thm2}{Theorem}
\newtheorem{lem2}{Lemma}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Classless Java}
\subtitle{Interface Based Programming for the Masses}

\authorinfo{}
           {}
           {}
%\authorinfo{Name2\and Name3}
%           {Affiliation2/3}
%           {Email2/3}

\maketitle

\begin{abstract}

  This paper presents Interface Based (IB) object-oriented (OO)
  programming. Unlike class based OO programming, in IB programming there
  is no need for classes. To support state, IB programming does not
  use fields directly. Instead it uses \emph{abstract state
    operations}, such as getters, setters or clone-like methods. This
  new way to deal with state allows flexibility not typically
  available in CB languages.  In IB state (including mutable state)
  can be automatically type-refined in subtypes. The combination of a
  purely IB style of programming and automatic type-refinement enables
  powerful idioms using multiple inheritance and state.  To introduce
  IB to programmers we created Classless Java: an embeding of IB
  directly into Java. Classless Java uses annotation processing for
  code generation and relies on the new features of Java 8 for
  interfaces. We illustrate the usefulness of IB through various
  examples and case studies.

\end{abstract}

\begin{comment}
\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2
\end{comment}

\input{sections/Introduction.tex}

\input{sections/Overview.tex}

\input{sections/Implementation.tex}

\input{sections/Formalization.tex}

\input{sections/Translation.tex}

\input{sections/CaseStudy.tex}

\input{sections/FutureWork.tex}

\input{sections/RelatedWork.tex}

\input{sections/Conclusion.tex}

%\acks
%
% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{paper}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\appendix
\input{sections/Appendix_Formalization.tex}
\input{sections/Appendix_Translation.tex}
\input{sections/Appendix.tex}

\end{document}