### 2017-03-31 FACS/Extensions.tex

\section{Extensions to our model}
  One of the main feature of our simple reuse/use model is that it can be
  easly extended. One simple but amazingly expressive extension is nested classes

\subsection{Nested classes}

A nested class will be another kind of member in the Literal, so  
the grammar can be updated as following:

\begin{bnf}
\prodFull\mMD{
\Opt{\terminalCode{class}}\ \terminalCode{method}\ \mT\ \mm\oR\overline{\mT\,\mx}\cR \Opt\me
\mid \mCD
}{Member Decl}\\
\prodFull\mT{\mC\mid\mC\terminalCode{.}\mT}{types are now paths}
\end{bnf}\\

The general idea is that by composing code with \use, nested classes with the same name are recursivelly composed.
Note that while we have nested classes, we do not have nested traits: all traits are still
at top level.
Untypable/unresolved Traits are also the only``dependency"
the type system keeps track of, this means that when a nested class at an arbirary
nested level is flattend, as in
\Q@C:{ D:{ E:Use t1,t2,L}}@
t1 and t2 must be defined before C at top level; and they may require classes (and their
nested) defined before C. This means that the type system can still consider
the class table as a simple map from Types T to their definition.

This extension lets us challenge the expression problem[]:
in the expression problem we have data variants and operations and....

Let see how to easily encode and solve the expression problem:

\begin{lstlisting}
exp:{Exp:{interface}}//Exp declared once, reused everywhere
lit:Use exp,{ Lit:{implements Exp //Exp not explicitly declared
    class method Lit of(int inner) //Lit abstract state
    method int inner()}
  }
sum:Use  exp,{ Sum:{implements Exp 
    class method This of(Exp left, Exp right)
    method Exp left() //Sum abstract state
    method Exp right()}
  }  
uminus:Use exp,{ UMinus:{implements Exp 
    class method This of(exp inner)//and so on for
    method Exp inner()}//all the needed datavariants
  }

expToS:{Exp:{interface method String toString()}}
//concept of toString declared once

sumToS:Use sum,expToS,{ Sum:{implements Exp//with Exp.toString
    method String toString()//just the implementation of the
      left.toString()+"+"+right.toString()//specific method
  }
uminusToS:...//implement toString for all the datavariants

expEval:{Exp:{interface method int eval()}}
//declare the next operation and implement it for all the datavariant
\end{lstlisting}

Now that you have nicely modularized the code, just compose all the traits you need.
\Q@MySolution:Use sumToS,litToS@ //sum,lit and exp traits are alread included

The expression problem presented up to now is the traditional challenge proposed by [];
this has been criticized to not really address the fundamental issues since it does not handle ....
Now we show how we can go behond the traditional expression problem by encoding transformer methods:
For example, lets add 1 to all literals
\begin{lstlisting}
expAdd1:{Exp:{interface method Exp add1()}}
sumAdd1:Use sum,expAdd1,{Sum:{implements Exp
    method Exp add1()
      Sum.of(left.add1,right.add1())
  }
litAdd1:Use lit,expAdd1,{Lit:{implements Exp
    method Exp add1()
      Lit.of(inner()+1);
    }

MySolutionAdd1:Use sumToS,litToS,sumAdd1,litAdd1
\end{lstlisting}

This nicely solve our problem. 
However, notice how if we wished to add many similar operations we would 
have to repeat the propagation code (as in \Q@sumAdd1@) many times
just changing the name of the operation.
In the next section we will show how to improve on this point.

\section{More composition operators}
\use is amazing, elegant and simple, but the system can be easily enriched with more 
operators: while most approaches in literator present a fixed set of operators, in our
system we do not need such restriction.
We just need to be sure that every newly added operator respects the following criteria:
if starts form well typed code, it produces well typed code.\marco{not enough? rephrease to respect the two quality at the start?}

---------
Restrict make abstract method

--------
Rename, like refactoring
Rename can invoke sum inside
---------
Redirect, like generics
Redirect can be multiple.

With redirect+rename we can have the general operator propagator

operation:{//for sum and lit, easy to extends as before
  T:{}
  Exp:{interface method Exp op(T x)}}
  Sum:Sum sum,{ extends Exp sum,expAdd1,{
    method Exp op(T x)
      Sum.of(left.op(x),right.op(x));
    }
  Lit:Sum lit,{
    method Exp op(T x)
      this;
  }
Now, to have my addN I can
opAddn: Sum
  operation[redirect T to Int]
    [rename Exp.op(x) to addN(x)][restrict Lit.op(x)], {
  Lit:{method Exp addN(Int x) Lit.of(inner())+x}
  }

Full power of redirect

serviceCombinator:{ S:{implements Service
  method R report()
  }

  R:{method R combine(R that)
    class method R empty()}

  ListS:list[redirect Elem to S]

  class R method(ListS ss){//here we use extended java like syntax
    R r=R.empty()
    for(S s in ss){
      s.performService();
      r=r.combine(s.report())
      }
    return r;
  }
}
PaintingService:serviceCombinator[redirect S to Painting.Service]

To encode the former generic code in java you need to write
the following headeche inducing interfaces for RService and Report (and still we miss the empty method)
and require that the services you want to serve implement those.
  interface Service{ void performService();}
  interface Report<R extends Report<R>>{R combine(R that);}
  interface RService<R extends Report<R>> extends Service{ R report();}