### 2017-05-19 FACS/section2.tex
.

\section{The Design of \name: Separating Use and Reuse}

This section presents the overview of \name, and it illustrates the
key ideas of its design. In particular we illustrate how to separate code use and 
code reuse in \name. %, and how this allows improving both.

%\subsection{The Design of \name: Decoupling Use from Reuse}

\subsection{Classes in \name: A mechanism for code Use}
%\name has a substantially different design from Java-like languages.
The concept of a class in \name provides a mechanism for code use
only. This means that in \name there is actually no subclassing, and
classes are roughly equivalent to final classes in Java.  Thus,
compared to Java-like languages, the most noticeable difference is the
absence of the \Q@extends@ keyword in \name.

To illustrate classes in \name consider the example in Section~\ref{}:

\begin{lstlisting}
 A: { method int ma() Utils.m(this) }//note, no {return _}
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

\noindent Classes in \name use a slightly different declaration style compared
to Java. In \name there is no \lstinline{class} keyword. 
Class declarations have a name (which must always start with
an uppercase letter) and a code literal, which is used to specify the
definitions of the class. For instance, in the class declaration
for \lstinline{A}, the name of the class is \lstinline{A} and the code 
literal associatted with the class (\Q@{ method int ma()   Utils.m(this) }@) contains the definitions associated to the
class. There are, however, some 
important differences to Java-like languages in the way 
classes and code-literals are type-checked, as we shall see next. 
Nevertheless, for this example, things still work in a similar way to Java.

The \name code above is fine, but there is no way to add a class 
\Q@B@ reusing the code of \Q@A@, since
\Q@A@ is designed for code \emph{use} and not \emph{reuse}. So, unlike
the Java code, introducing a subclass
\lstinline{B} is not possible. At first, this may seem like a severe restriction, but
\name has a different mechanism for code \emph{reuse} that 
is more appropriate when code reuse is intended.

\subsection{Traits in \name: A mechanim of code Reuse}

Unlike classes traits in \name cannot be instantiated and do not introduce new
types. However they provide code reuse.
%So, lets try again encoding the code for the leaking problem, but this
%time aiming at code reuse. 
Trait declarations 
look very much like class declarations, but trait names 
start with a lowercase letter. An obvious first attempt 
to model the example in Section~\ref{} (for a Java programmer) 
with traits and code reuse is:

\begin{lstlisting}
 ta:{ method int ma() Utils.m(this) }//type error
 A:Use ta
 Utils:{ class method int m(A a)/*method body here*/ }
\end{lstlisting}

\noindent Here \lstinline{ta} is a trait intended to replace the
original class \lstinline{A} so that the code of the method
\lstinline{ma} can be reused. Then the class \lstinline{A} 
is created by inheriting the code from the trait using the keyword 
{\bf Use}. Note that \use\ cannot contain class names: only trait
names are allowed.
That is, using a trait is the only way to induce code reuse.
Unfortunatelly, this code does not work, 
because \lstinline{Utils.m} requires an \lstinline{A} and the type of \lstinline{this} in
\Q@ta@ has no relationship to the type \lstinline{A}.
Since the trait name is not a type, no code external to that trait can
refer to it. This is one of the key design decisions in \name. 
The point is that the trait can refer to the program via \lstinline{this}, but the
program is agnostic to what the trait type is going to be, so it can
be later assigned to any (or many) classes. This improves the
flexibility of reuse as illustrated in Section~\ref{}. However, 
to solve the more immediate typing issue above we need one more
round of refactoring, as we shall see next.

%: in \name all declarations 
%are sequentially type-checked and if a definition occurs afterwards
%another, then it is not visible on all previous definitions.

\paragraph{Type of the self-reference} The code above does not work
because the way type-checking works in \name is that type-checking of 
code literals is independent of the class/trait names associated to
it. A Java programmer may expect that the type of \lstinline{this} 
in the previous definition of \lstinline{ta} is \lstinline{ta}
itself. However, this intuition brought from Java is wrong in \name
for two reasons:

\begin{itemize}

\item The first reason is that {\bf traits are not types}. Traits in
  \name are simply units of reuse and cannot be used as types. Types
  are only introduced by class/interface declarations.

\item The second reason is that {\bf the type of self-references is the self type of the
    code literal.} Unlike Java-like languages, where the body of a
  class declaration is intrinsically coupled with the class itself, in
  \name code literals are first-class\bruno{Is it appropriate to say
    that they are first class?} and are type-checked independently.
  \name has a notion of self-type, which is closely related to
  approaches such as ThisType~\cite{} and ...\cite{others}. Therefore
  in \name the self-reference does not have the type of the class
  being defined, but rather it has the self-type.
\end{itemize}

With this in mind, we can try to model the example in Section~\ref{} again:

\begin{lstlisting}
 IA:{interface method int ma()}//interface with abstract method
 Utils:{ class method int m(IA a)/*method body here*/}
 ta:{implements IA
   method int ma() Utils.m(this) }
 A:Use ta
\end{lstlisting}

This code works: \Q@Utils@ is using an interface \Q@IA@ and the trait \Q@ta@
is implementing it. Thus the self-type of the trait \Q@ta@ (and the
type of \lstinline{this}) will be a
subtype of \Q@IA@. It is also possible to add a \Q@B@ as follows
\begin{lstlisting}
  B:Use ta, { method int mb(){return this.ma();} }
\end{lstlisting}
This also works.  \Q@B@ reuses the code of \Q@ta@, but has no knowledge of \Q@A@.
Since \Q@B@ reuses \Q@ta@, and \Q@ta@ implements \Q@IA@, also \Q@B@ implements \Q@IA@.

\paragraph*{Semantic of Use}
Albeit alternative semantic models for traits~\cite{} have been proposed,
here we use the flattening model.
This means that 
\begin{lstlisting}
A:Use ta
B:Use ta, { method int mb(){return this.ma();} }
\end{lstlisting}

\noindent reduces/is equivalent to/is flatted into

 \begin{lstlisting}
A:{implements IA method int ma() Utils.m(this) }
B:{implements IA
  method int ma() Utils.m(this)
  method int mb() this.ma() }

 \end{lstlisting}

 This code seems correct, and there is no mention of the trait
 \Q@ta@. In some sense, all the information about code
 reuse/subclassing is just a private implementation detail of \Q@A@
 and \Q@B@; while subtyping is part of the class interface.

\begin{figure}[t]
\begin{center}
\begin{tabular}{c|c|c|c}
              & Instantiable (Use) & Unit of Reuse & Introduces New Type \\
\hline
Class  & Yes              & No          & Yes           \\
Trait   & No               & Yes         & No            \\
\end{tabular}
\end{center}
More complete table here, we can put it where you prefer
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c}
&\Rotated{direct instantation}
&\Rotated{self instantiation}
&\Rotated{unit of use}
&\Rotated{unit of reuse}
&\Rotated{introduce type}
&\Rotated{induced type is this type}
&\Rotated{binary methods}
&\Rotated{reuse induce subtype}
&\Rotated{well-typed before imported}
&\Rotated{well-typed before composed} 
\\
\hline
java/scala class&3Y&X&Y&Y&Y&Y&X&Y&Y&X\\
java8 interface &X&X&X&Y&Y&Y&X&Y&Y&X\\
scala trait        &X&X&X&Y&Y&Y&X&Y&X&X\\
original trait     &X&X&X&Y&-&-&X&X&-&-\\
Ferruccio trait  &X&X&X&Y&X&-&X&X&Y&Y\\
42 trait            &X&Y&X&Y&X&-&Y&X&Y&X\\
42 class           &Y&Y&Y&X&Y&Y&Y&-&Y&X\\
module composition
                      &-&-&Y&Y&-&-&-&-&Y&Y\\
deepFJig class &Y&Y&Y&Y&Y&X&Y&X&Y&Y\\
package template
                      &X&Y&X&Y&X&-&X&X&Y&X\\

\end{tabular}
\end{center}

\caption{A comparision between traits and classes in \name.}
\label{fig:compare}
\end{figure}

To finish this section, Figure \ref{fig:compare} provides a summary of
the differences between classes and traits. The comparison focus on
the roles of traits and classes with respect to instantiation,
reusability and whether the declarations also introduce new types or
not.